### 2017-11-10 paper/sections/1_introduction.tex
paper: addressed #8
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

% What is Mantle
We use the data management language and policy engine from the Mantle
paper~\cite{sevilla:sc15-mantle} to dynamically explore the effects of
different cache management strategies for the changing key-value workloads
generated by ParSplice.  Mantle was originally touted as a programmable file
system metadata load balancer, but we realize now that the collection of
abstractions designed for file systems was a control plane that improved
metadata access. So in this paper we refer to Mantle as a policy engine that
injects policies written in our data management language directly into a
running service ({\it e.g.} file system, key-value store, application) and show
how this approach is useful for reasoning about and designing different cache
management strategies in ParSplice. Developers write policies for ``when" they
want data moved and ``how much" of the data to move, then the framework
executes these policies whenever a decision needs to be made.  These
abstraction help developers unfamiliar with the domain quickly reason about,
develop, and deploy new policies that control temporal and spatial locality. We
show that Mantle:

>>>>

% What is Mantle
We use the data management language and policy engine from the Mantle
paper~\cite{sevilla:sc15-mantle} to dynamically explore the effects of
different cache management strategies for the changing key-value workloads
generated by ParSplice.  Mantle was originally touted as a programmable file
system metadata load balancer, but we realize now that the collection of
abstractions designed for file systems was a control plane that improved
metadata access. So in this paper we refer to Mantle as a policy engine that
injects policies written in our data management language directly into a
running service and show how this approach is useful for reasoning about and
designing different cache management strategies in ParSplice.  By service, we
mean a system that manages data and responds requests, such as a file system or
key-value store.  Developers write policies for ``when" they want data moved
and ``how much" of the data to move, then the framework executes these policies
whenever a decision needs to be made.  These abstraction help developers
unfamiliar with the domain quickly reason about, develop, and deploy new
policies that control temporal and spatial locality. We show that Mantle:

### 2017-11-10 paper/sections/4_dynamic-policies.tex
paper: addressed #8
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

% results: raw numbers
Although the keyspace grows to 150K, a 100K key cache achieves 99\% of the
performance. Decreasing the cache degrades performance and predictability.
While this result is not unexpected, it nonetheless achieves our goal of
showing the benefits of load balancing keys across nodes and that smaller
caches on each node are an effective way to save memory without completely
sacrificing performance.

% why is the performance lower for smaller caches?
Despite the memory savings, our results suggest that dynamic load balancing
policies could save even more memory.  Figure~\ref{fig:methodology-tradeoff}
show that a 100K key cache is sufficient as a static policy but the top graph
in Figure~\ref{fig:motivation-regimes} indicates that the cache size could be
much smaller. That graph shows that the beginning of the run is characterized
by many reads to a small set of keys and the end sees much lower reads per
second to a larger keyspace. Specifically, it shows only about 100 keys as
active in the latter half of the run.

After analyzing traces, we see that the 100 key cache is insufficient because
the persistent database cannot service the read-write traffic.  According to
Figure~\ref{fig:motivation-regimes}, the read requests arrive at 750 reads per
second in addition to the writes that land in each tier (about 300 puts/second,
some redundant). This traffic triggers a LevelDB compaction and reads block,
resulting in very slow progress.  Traces verify this hypothesis and show reads
getting backed up as the read/write ratio increases. To recap, small caches
incur too much load on the persistent database  at the beginning of the run but
should suffice after the initial read flash crowd passes because
the keyspace is far less active. This suggests a two-part load balancing
policy.

@@

\begin{figure}[t]
  \noindent\includegraphics[width=0.5\textwidth]{figures/methodology-tradeoff-dynamic.png}\\
  \caption{The performance and resource utilization trade-off of using a
  dynamic load balancing policy that switches to a constrained cache policy after absorbing
  the initial burstiness of the workload. The sizes of these smaller caches are
  on the \(x\) axis.  \label{fig:methodology-tradeoff-dynamic}}
\end{figure}

Figure~\ref{fig:methodology-tradeoff-dynamic} shows the results of using Mantle
to program a dynamic load balancing policy into ParSplice that switches between
two policies:

@@

The results in Figure~\ref{fig:methodology-tradeoff-dynamic} are slightly
deceiving for three reason: (1) segments take longer to generate later in the
run, (2) the memory footprint is the value at the end of 2.5 hours, and (3)
this policy only works well for the 2.5 hour run.  For (1), the curving down of
the simulation vs. wall-clock time is shown in
Figure~\ref{fig:methodology-trajectory}; as the nanoparticle grows it takes
longer to generate segments so by the time we reach 2 hours, over 90\% of the
trajectory is already generated.  For (2), the memory footprint is around 0.4GB
until we reach 100K key switch threshold. In
Figures~\ref{fig:methodology-tradeoff}
and~\ref{fig:methodology-tradeoff-dynamic} we plot the final value. For (3),
Figure~\ref{fig:methodology-trajectory} shows that the cache fills up with 100K
keys at time 7200 seconds and its size is reduced to the size listed in the
legend.  The curves stay close to ``Unlimited" for up to an hour after the
cache is reduced but eventually flatten out as the persistent database gets
overloaded. 10K and 100K follow the ``Unlimited" curve the longest and are
sufficient policies for the 2.5 hour runs but anything longer would need a
different dynamic load balancing policy.

@@

% but the result is still valid
Despite these caveats, the result is still valid: we found a dynamic load
balancing policy that absorbs the cost of a high read throughput on a small
keyspace and reduces the memory pressure for a 2.5 hour run. Our experiments
show the effectiveness of the load balancing policy engine we integrated into
ParSplice, not that we were able to identify the best policy for all system
setups ({\it i.e.} different ParSplice parameters, number of worker tasks, and
job lengths).  To solve that problem, we need a way to identify what thresholds
we should use for different job permutations.

>>>>

% why is the performance lower for smaller caches?
Although the keyspace grows to 150K, a 100K key cache achieves 99\% of the
performance. Decreasing the cache degrades performance and predictability.
Despite the memory savings, our results suggest that dynamic cache management
policies could save even more memory.  Figure~\ref{fig:methodology-tradeoff}
show that a 100K key cache is sufficient as a static policy but the top graph
in Figure~\ref{fig:motivation-regimes} indicates that the cache size could be
much smaller. That graph shows that the beginning of the run is characterized
by many reads to a small set of keys and the end sees much lower reads per
second to a larger keyspace. Specifically, it shows only about 100 keys as
active in the latter half of the run.

After analyzing traces, we see that the 100 key cache is insufficient because
the persistent database cannot service the read-write traffic.  According to
Figure~\ref{fig:motivation-regimes}, the read requests arrive at 750 reads per
second in addition to the writes that land in each tier (about 300 puts/second,
some redundant). This traffic triggers a LevelDB compaction and reads block,
resulting in very slow progress.  Traces verify this hypothesis and show reads
getting backed up as the read/write ratio increases. To recap, small caches
incur too much load on the persistent database  at the beginning of the run but
should suffice after the initial read flash crowd passes because
the keyspace is far less active. This suggests a two-part cache management
policy.

@@

\begin{figure}[t]
  \noindent\includegraphics[width=0.5\textwidth]{figures/methodology-tradeoff-dynamic.png}\\
  \caption{The performance and resource utilization trade-off of using a
  dynamic cache management policy that switches to a constrained cache policy after absorbing
  the initial burstiness of the workload. The sizes of these smaller caches are
  on the \(x\) axis.  \label{fig:methodology-tradeoff-dynamic}}
\end{figure}

Figure~\ref{fig:methodology-tradeoff-dynamic} shows the results of using Mantle
to program a dynamic cache management policy into ParSplice that switches between
two policies:

@@

The results in Figure~\ref{fig:methodology-tradeoff-dynamic} are slightly
deceiving for three reason: (1) segments take longer to generate later in the
run, (2) the memory footprint is the value at the end of 2.5 hours, and (3)
this policy only works well for the 2.5 hour run.  For (1), the curving down of
the simulation vs. wall-clock time is shown in
Figure~\ref{fig:methodology-trajectory}; as the nanoparticle grows it takes
longer to generate segments so by the time we reach 2 hours, over 90\% of the
trajectory is already generated.  For (2), the memory footprint is around 0.4GB
until we reach 100K key switch threshold. In
Figures~\ref{fig:methodology-tradeoff}
and~\ref{fig:methodology-tradeoff-dynamic} we plot the final value. For (3),
Figure~\ref{fig:methodology-trajectory} shows that the cache fills up with 100K
keys at time 7200 seconds and its size is reduced to the size listed in the
legend.  The curves stay close to ``Unlimited" for up to an hour after the
cache is reduced but eventually flatten out as the persistent database gets
overloaded. 10K and 100K follow the ``Unlimited" curve the longest and are
sufficient policies for the 2.5 hour runs but anything longer would need a
different dynamic cache management policy.

@@

% but the result is still valid
Despite these caveats, the result is still valid: we found a dynamic cache 
management policy that absorbs the cost of a high read throughput on a small
keyspace and reduces the memory pressure for a 2.5 hour run. Our experiments
show the effectiveness of the policy engine we integrated into
ParSplice, not that we were able to identify the best policy for all system
setups ({\it i.e.} different ParSplice parameters, number of worker tasks, and
job lengths).  To solve that problem, we need a way to identify what thresholds
we should use for different job permutations.
